---
title: "Joe_Rovalino_wk#3_math"
author: "Joe Rovalino"
date: "8/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) Data Exploration: This should include summary statistics, means, medians, quartiles, or any
other relevant information about the data set. Please include some conclusions in the R
Markdown text.

 
```{r}
#install.packages("ggplot2")
require(ggplot2)

#Loads the data from URL. Credited data creator in github description. Basic # of rows and Columns. This is S&P Stock data

  urldata <- read.csv("https://raw.githubusercontent.com/jrovalino/RWk3FinalProject/master/Smarket.csv", header = TRUE, sep = ",")
#urldata
nrow(urldata)
ncol(urldata)
str(urldata)
```
#Getting stats on the data and performing data exploration:

```{r}
summary(urldata)
#Mean and Median for "Volumes" vector in the data.frame urldata for Volume Column
YrsofData <-unique(urldata$Year)
YrsofData
data2001 <- subset(urldata, Year == 2001)
#data2001
#2001 Mean and Median
Vol2001MeanRow <- mean((data2001[["Volume"]]))
Vol2001MeanRow
Vol2001MedianRow<-median(c(data2001[,8]))
Vol2001MedianRow
data2002 <- subset(urldata, Year == 2002)
#data2002
#2002 Mean and Median
Vol2002MeanRow <- mean((data2002[["Volume"]]))
Vol2002MeanRow
Vol2002MedianRow<-median(c(data2002[,8]))
Vol2002MedianRow
data2003 <- subset(urldata, Year == 2003)
#data2003
#2003 Mean and Median
Vol2003MeanRow <- mean((data2003[["Volume"]]))
Vol2003MeanRow
Vol2003MedianRow<-median(c(data2003[,8]))
Vol2003MedianRow
data2004 <- subset(urldata, Year == 2004)
#data2004
#2004 Mean and Median
Vol2004MeanRow <- mean((data2004[["Volume"]]))
Vol2004MeanRow
Vol2004MedianRow<-median(c(data2004[,8]))
Vol2004MedianRow
data2005 <- subset(urldata, Year == 2005)
#data2005
#2005 Mean and Median
Vol2005MeanRow <- mean((data2005[["Volume"]]))
Vol2005MeanRow
Vol2005MedianRow<-median(c(data2005[,8]))
Vol2005MedianRow
#Total Mean and Median
TotVolMeanRow <- mean((urldata[["Volume"]]))
TotVolMeanRow
TotVolMedianRow<-median(c(urldata[,8]))
TotVolMedianRow
Perc2005Better<-(Vol2005MeanRow/TotVolMeanRow)
Perc2005Better
Perc2001Worse<-1-(Vol2001MeanRow/TotVolMeanRow)
Perc2001Worse
```
#Volume Analysis:
The Volume mean across the years(2001-2005 respectively): 1.22695, 1.428451, 1.384693, 1.424159, 1.917298
Total Volume mean: 1.478305
The volume mean in 2005 was the highest and 29% higher than the total mean (2001-2005). 
The volume mean in 2001 was the lowest and 17% lower than the total mean (2001-2005).


2) Data wrangling: Please perform some basic transformations. They will need to make sense but
could include column renaming, creating a subset of the data, replacing values, or creating new
columns with derived data (for example – if it makes sense you could sum two columns
together)
```{r}
colnames(urldata)<-c("X","Year","%1D", "%2D","%3D", "%4D", "%5D","Vol","%Today","Dir")
head(urldata)
tail(urldata)
```
Change to columns name from Lag# columns to more meaningful name that represents the percent return for # Day prior. ex: %1D.

Example of data wrangling is also demonstrated by the creation of 5 subsets by year for new dataframes used above and also renaming the lag fields.
```{r}
colnames(data2001)<-c("X","Year","%1D", "%2D","%3D", "%4D", "%5D","Vol","%Today","Dir")
head(data2001)
colnames(data2002)<-c("X","Year","%1D", "%2D","%3D", "%4D", "%5D","Vol","%Today","Dir")
head(data2002)
colnames(data2003)<-c("X","Year","%1D", "%2D","%3D", "%4D", "%5D","Vol","%Today","Dir")
head(data2003)
colnames(data2004)<-c("X","Year","%1D", "%2D","%3D", "%4D", "%5D","Vol","%Today","Dir")
head(data2004)
colnames(data2005)<-c("X","Year","%1D", "%2D","%3D", "%4D", "%5D","Vol","%Today","Dir")
head(data2005)
```
3) Graphics: Please make sure to display at least one scatter plot, box plot and histogram. Don’t
be limited to this. Please explore the many other options in R packages such as ggplot2.

```{r}
boxplot(data2001$'%Today')
hist(data2001$'%Today')
qplot(Vol, '%Today', data=data2001)
boxplot(data2002$'%Today')
hist(data2002$'%Today')
qplot(Vol, '%Today', data=data2002)
boxplot(data2003$'%Today')
hist(data2003$'%Today')
qplot(Vol, '%Today', data=data2003)
boxplot(data2004$'%Today')
hist(data2004$'%Today')
qplot(Vol, '%Today', data=data2004)
boxplot(data2005$'%Today')
hist(data2005$'%Today')
qplot(Vol, '%Today', data=data2005)
boxplot(urldata$'%Today')
hist(urldata$'%Today')
qplot(Vol, '%Today', data=urldata)
```

4) Meaningful question for analysis: Please state at the beginning a meaningful question for
analysis. Use the first three steps and anything else that would be helpful to answer the
question you are posing from the data set you chose. Please write a brief conclusion paragraph
in R markdown at the end.

What was the largest Percent day change- Up and Down? Which year did it occur? Visually display %today? (Hint graph)

```{r}
str(urldata)
urldata[which.max(urldata$'%Today'),]
urldata[which.min(urldata$'%Today'),]
#urldata
hist(urldata$'%Today')
```

Answer: 

The largest up % change was 5.733 in 2002

The largest down % was -4.922 in 2001.

5) BONUS – place the original .csv in a github file and have R read from the link. This will be a very
useful skill as you progress in your data science education and career.

Answer: 
Please see beginning of code - it is pulling from github using:
  urldata <- read.csv("https://raw.githubusercontent.com/jrovalino/RWk3FinalProject/master/Smarket.csv", header = TRUE, sep = ",")
